---
title: "California Schools by Zip Code"
author: "Ryan Estrellado"
date: "5/26/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Purpose 

The purpose of this Shiny App is to take a California five digit zip code and 
plot a map that shows California public schools in that area. Some possible 
use cases for this app are: 

- Parents who are looking to move to California or within California can search 
for schools that are in the area they are interested in
- Analysts who are researching California schools can have quick access to 
the names of schools in a given area 

## About the Data 

The data used for this app is publicly available and can be found at the 
[California Department of Education's website](http://www.cde.ca.gov/ds/si/ds/pubschls.asp). The data has many 
useful variables, including longitude and latitude. 

## About the Data (cont)

The app draws on a rich dataset that includes all California districts and 
schools. 

```{r message = F, warning = F, comment = NULL}
library(tidyverse)
library(knitr)

# Read and clean data
  if(file.exists("schools_by_zip/data/pubschls.txt") == FALSE) {
    download.file("ftp://ftp.cde.ca.gov/demo/schlname/pubschls.txt",
                  "data/pubschls.txt",
                  method = "curl")
  }
  
  original <- read_tsv("schools_by_zip/data/pubschls.txt") 
```

```{r echo = TRUE}
kable(
  original %>% 
  filter(StatusType == "Active") %>% 
  summarise(`total schools` = n_distinct(School), 
            `total districts` = n_distinct(District))
)
```

## Server Calculation (Zip Code 95670)

```{r message = F}
library(leaflet)
library(stringr)

cde_schls <- original %>%
      filter(StatusType == "Active",
             str_detect(MailZip, "95670") == TRUE) %>%
      rename(lat = Latitude, lon = Longitude) %>%
      select(-c(AdmFName1:AdmEmail3, FundingType, Magnet, StatusType,
                Street:State, DOC, CharterNum, SOC, EdOpsCode, EILCode))
```

```{r echo = T, message = F, fig.width = 5, fig.height = 3}
# Takes a zip code as input and generates a map
cde_schls %>% 
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(popup = cde_schls$School)
```


